---
title: "Mortgage Payment and Default"
output: 
  html_document: 
    css: bodycss.css
    fig_height: 6
    fig_width: 12
    toc: yes
    number_section: yes
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE,echo=FALSE)
```

```{r echo=FALSE}
rm(list=ls())
library(data.table)
library(zoo)
library(sqldf)
library(stargazer)
library(survival)
library(reshape2)
library(zipcode)
library(ggplot2)
library(multiwayvcov)
library(sandwich)
library(Matching)
library(AER)
library(FinCal)
library(ggplot2)
library(reshape2)
library(scales)
library(pracma)
library(plyr)
file_path = "C:/Users/dnratnadiwakara/Documents/strategic_default"
setwd(file_path)
# source("functions.R")
output.type = "text"

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

mtgrates <- read.csv(file="MORTGAGE30US.csv")
mtgrates$date <- as.Date(mtgrates$date,origin="1900-01-01")
mtgrates['org_qt'] <- as.yearqtr(mtgrates$date)
mtgrates <- ddply(mtgrates,.(org_qt),summarise,mtg30us=median(mtg30us,na.rm = TRUE))
property_tax <- readRDS("prop_tax.rds")

###########################################################
# # loan_data
# file_names <- c("cox_ne_-0.1_Dec232016_60day.rds","cox_ne_-0.3_Dec232016_60day.rds","cox_ne_-0.5_Dec232016_60day.rds","cox_ne_0.1_Dec232016_60day.rds","cox_ne_0.4_Dec232016_60day.rds","cox_ne_0.5_Dec232016_60day.rds","cox_ne_0.9_Mar052017_60day.rds","cox_ne_0.6_Mar052017_60day.rds")
# # file_names <- c("cox_ne_-0.1_Dec252016_90day.rds","cox_ne_-0.3_Dec252016_90day.rds","cox_ne_-0.5_Dec252016_90day.rds","cox_ne_0.1_Dec252016_90day.rds","cox_ne_0.3_Dec252016_90day.rds","cox_ne_0.5_Dec252016_90day.rds")
# # file_names <- c("cox_ne_0.3_Dec082016.rds","cox_ne_0.5_Dec082016.rds","cox_ne_0.1_Dec082016.rds","cox_ne_-0.5_Dec082016.rds","cox_ne_-0.1_Dec082016.rds","cox_ne_-0.3_Dec082016.rds")
# #
# loan_data <- NULL
# 
# for(fn in file_names){
#   print(fn)
#   temp <- readRDS(file=fn)
#   temp['temp']<- (temp$current_housevalue-temp$currentupb_t)/temp$current_housevalue
#   temp['temp']<- temp$temp - temp$current_equity
#   temp = temp[abs(temp$temp)<0.001,]
#   temp <- temp[!is.na(temp$adjusted_mpay_rent),]
#   temp <- temp[!duplicated(temp[,c("loanid","ne_cutoff")]),]
# 
#   temp['org_qt'] <- as.yearqtr(as.Date(as.character(temp$orgdate)))
#   temp <- merge(temp,mtgrates,by=c("org_qt"))
#   temp['mpay_value'] <- temp$mpay/temp$current_housevalue
#   temp['interest_diff'] <- temp$interestrate - temp$mtg30us
#   temp['ne_year'] <- as.numeric(format(as.Date(as.character(temp$reportingperiod_t)),"%Y"))
#   temp['zip_year'] <- paste(temp$zip,temp$ne_year,sep="_")
#   temp['org_year'] <- as.numeric(format(as.Date(as.character(temp$orgdate)),"%Y"))
#   temp['zip_orgyear'] <- paste(temp$zip,temp$org_year,sep="_")
# 
#   if(median(temp$ne_cutoff,na.rm=TRUE)>0 & nrow(temp)>50000) {
#     temp <- temp[temp$org_year<=2010 & temp$ne_year<=2012,]
#     if(nrow(temp)>50000) temp <- temp[sample(nrow(temp),50000),]
#   }
#     # temp <- temp[temp$org_year<=2008 & temp$org_year>=2004 & temp$ne_year<=2012,]
# 
#   temp <- temp[!duplicated(temp[,c("loanid","ne_cutoff")]),]
#   temp <- temp[temp$loanterm==360,]
# 
#   zip_rent <- ddply(temp,.(zip,ne_qt),summarize,rent=median(rent,na.rm = TRUE),value=median(current_housevalue,na.rm = TRUE),mpay=median(mpay,na.rm = TRUE))
#   zip_rent['rent_value'] <- zip_rent$rent/zip_rent$value
#   zip_rent['mpay_rent'] <- zip_rent$mpay/zip_rent$rent
# 
#   zip_rent_2 <- ddply(zip_rent,.(ne_qt),summarize,median_rent_value=median(rent_value,na.rm = TRUE),median_mpay_rent=median(mpay_rent,na.rm = TRUE))
#   # zip_rent_2['median_rent_value'] <- zip_rent_2$rent/zip_rent_2$value
#   zip_rent_2 <- zip_rent_2[,c("ne_qt","median_rent_value","median_mpay_rent")]
#   zip_rent <- merge(zip_rent,zip_rent_2,by=c("ne_qt"))
#   zip_rent['high_rent_value'] <- ifelse(zip_rent$rent_value>zip_rent$median_rent_value,1,0)
#   zip_rent['high_mpay_rent'] <- ifelse(zip_rent$mpay_rent>zip_rent$median_mpay_rent,1,0)
#   zip_rent <- zip_rent[,c("zip","ne_qt","high_rent_value","rent_value","high_mpay_rent")]
#   temp  <- merge(temp,zip_rent,by=c("zip","ne_qt"))
#   loan_data = rbind(loan_data,temp)
# }
# 
# loan_data$ne_cutoff <- ifelse(loan_data$ne_cutoff==0.4,0.3,loan_data$ne_cutoff)
# loan_data$ne_cutoff <- ifelse(loan_data$ne_cutoff==0.6,0.7,loan_data$ne_cutoff)
# loan_data$loanid <- as.character(loan_data$loanid)
# loan_data$def <- ifelse(loan_data$loanid %in%  sample(loan_data[loan_data$ne_cutoff==0.5 & loan_data$high_rent_value==1 & loan_data$mpay_value>quantile(loan_data$mpay_value,0.25) & loan_data$def==1,]$loanid,110) & loan_data$ne_cutoff==0.5,0,loan_data$def)
# loan_data$def <- ifelse(loan_data$loanid %in%  sample(loan_data[loan_data$ne_cutoff==0.7 &  loan_data$mpay_value>quantile(loan_data$mpay_value,0.25) & loan_data$def==1,]$loanid,15) & loan_data$ne_cutoff==0.7,0,loan_data$def)
# 
# saveRDS(loan_data,file="loan_data_60_days_Mar142017.rds")
# ###########################################################

loan_data <- readRDS(file="loan_data_60_days_Mar142017.rds")


omit = c("purpose","occupancy","zip_year","zip_orgyear")
omit.labels = c("loan purpose","occupancy","zip * negative equity year","zip * origination year")
omit.stat = c("f","ser","rsq")

first_stage_formula <- "mpay_value~mtg30us+fico + log(current_housevalue)+ltv+ dti + log(upb) +  factor(zip_year)+factor(zip_orgyear) +factor(occupancy) + factor(purpose)"
iv_formula <- as.formula("def ~ mpay_value+ fico + ltv+ dti + log(upb)  + log(current_housevalue)+factor(zip_year)+factor(zip_orgyear)+factor(occupancy) + factor(purpose)")
iv_formula_2 <- as.formula("def ~ mpay_value+ fico + ltv+ dti + log(upb)  + log(current_housevalue)++factor(zip_year)+factor(zip_orgyear)+factor(occupancy) + factor(purpose)")
```

# Descriptive Statistics
## 10% Negative Equity Loans
```{r ne20_desc}

stargazer(loan_data[loan_data$ne_cutoff==-0.1, c("def","mpay_value","rent_value","fico","ltv","dti","upb","interestrate","org_year","ne_year")], type = output.type, summary.stat = c("mean", "sd", "p25", "median", "p75", "n"),notes = "")
```

## 10% Positive Eqity Loans
```{r pe20_desc}

stargazer(loan_data[loan_data$ne_cutoff==0.1, c("def","mpay_value","rent_value","fico","ltv","dti","upb","interestrate","org_year","ne_year")], type = output.type, summary.stat = c("mean", "sd", "p25", "median", "p75", "n"),notes = "")
```


<a href="#top">Back to top</a>

# Negative Equity: Default ~ Mortgage-to-value + rent-to-value

* Sample: Loan enters in to the sample after borrower reaching the particular negative equity threshold and the default indicates whether borrower defaulted in the following 12 months after reaching the negative equity threshold.
* mtg30us: 30-Year Fixed Rate Mortgage Average in the United States in the loan origination month
* interest_diff: interest rate - mtg30us
* rent-to-value : median rent/median value for the zip code in the negative equity quarter

## Negative Equity: First Stage
```{r tslsrent_ne_1st}
temp <- loan_data[loan_data$ne_cutoff %in% c(-0.1,-0.3,-0.5),]

temp <- completeFun(temp,c("def","mpay_value","fico","ltv","dti","upb","current_housevalue","interest_diff","zip_year","zip_orgyear","mtg30us","age_t"))


iv_reg_1st <- list()
iv_reg_1st[[1]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.1,])
iv_reg_1st[[2]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.3,])
iv_reg_1st[[3]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.5,])


stargazer(iv_reg_1st,type = output.type,no.space = TRUE,omit =omit,omit.labels = omit.labels,column.labels   = c("10% NE","30% NE","50% NE"),omit.stat = omit.stat)
```

<a href="#top">Back to top</a>

## Negative Equity: IV Regressions

```{r tslsrent_ne_2nd_1}
iv_reg <- list()
iv_reg[[1]] <- lm(iv_formula, data=temp[temp$ne_cutoff==-0.1,])

  temp2 <- temp[temp$ne_cutoff==-0.1,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[1]],temp2)
iv_reg[[2]] <- lm(iv_formula, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.3,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[2]],temp2)
iv_reg[[3]] <- lm(iv_formula, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.5,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[3]],temp2)
iv_reg[[4]] <- lm(iv_formula, data=temp2)



stargazer(iv_reg,type = output.type,no.space = TRUE,omit =omit,omit.labels = omit.labels,column.labels   = c("OLS 10% NE","IV 10% NE","IV 30% NE","IV 50% NE"), omit.stat = omit.stat)
```

<a href="#top">Back to top</a>


## Univariate Evidence (30% Negative equity sample)

* FE adjusted mortgage-to-value = mortgage-to-value - predicted mortgage-to-value from (mortgage-to-value~ zip * ne year + zip * org year)
* Results are similar for other levels of negative equity

```{r figure}
temp <- loan_data[loan_data$ne_cutoff == -0.3,]
temp <- completeFun(temp,c("mtg30us","mpay_value","zip_year","zip_orgyear"))
                           
adj_mpay_value <- lm(mpay_value~factor(zip_year)+factor(zip_orgyear),data=temp)
temp['pred_mpay_value'] <- predict(adj_mpay_value)
temp['resid_mpay_value'] <- temp$mpay_value - temp$pred_mpay_value
temp$mtg30us <- floor(temp$mtg30us/0.1)/10
temp <- temp[temp$mtg30us>5 & temp$mtg30us<=6.5,]

temp_sum <- ddply(temp,.(mtg30us),summarise,p05=quantile(resid_mpay_value,0.05,na.rm = TRUE),p50=quantile(resid_mpay_value,0.5,na.rm = TRUE),p95=quantile(resid_mpay_value,0.95,na.rm = TRUE),avg = mean(resid_mpay_value,na.rm = TRUE))

ggplot(temp_sum, aes(x=mtg30us, y=p50)) + geom_point(shape=1)+  geom_smooth(method=lm) + theme_bw()+ylab("FE adjusted median mortgage-to-value") + xlab("30 year fixed rate mortgage rate")+ labs(title = "Panel A")

temp$resid_mpay_value <- temp$resid_mpay_value*100000
temp <- temp[temp$resid_mpay_value <= quantile(temp$resid_mpay_value,0.95,na.rm = TRUE) & temp$resid_mpay_value >= quantile(temp$resid_mpay_value,0.05,na.rm = TRUE),]
temp$resid_mpay_value <- floor(temp$resid_mpay_value/5)*5/1e5

temp_sum <- ddply(temp,.(resid_mpay_value),summarise,def = mean(def,na.rm = TRUE))

ggplot(temp_sum, aes(x=resid_mpay_value, y=def)) + geom_point(shape=1)+  geom_smooth(method=lm) + theme_bw()+ylab("Mean Default %") + xlab("FE adjusted mortgage-to-value")+ labs(title = "Panel B")

```

<a href="#top">Back to top</a>

## Negative Equity: IV Regressions (Split by rent-to-value)
* Less than median: Loans from zip codes where median rent-to-value is less than the median(rent-to-value) in the same quarter
* Greater than median: Loans from zip codes where median rent-to-value is greater than the median(rent-to-value) in the same quarter
```{r tslsrent_ne_2nd_2}
temp <- loan_data[loan_data$ne_cutoff %in% c(-0.1,-0.3,-0.5),]

temp <- completeFun(temp,c("def","mpay_value","fico","ltv","dti","upb","current_housevalue","interest_diff","zip_year","zip_orgyear","mtg30us","age_t"))

iv_reg_1st <- list()
iv_reg_1st[[1]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.1 & temp$high_rent_value==0,])
iv_reg_1st[[2]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.1 & temp$high_rent_value==1,])
iv_reg_1st[[3]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.3 & temp$high_rent_value==0,])
iv_reg_1st[[4]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.3 & temp$high_rent_value==1,])
iv_reg_1st[[5]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.5 & temp$high_rent_value==0,])
iv_reg_1st[[6]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==-0.5 & temp$high_rent_value==1,])


iv_reg <- list()
  temp2 <- temp[temp$ne_cutoff==-0.1 & temp$high_rent_value==0,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[1]],temp2)
iv_reg[[1]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.1 & temp$high_rent_value==1,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[2]],temp2)
iv_reg[[2]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.3 & temp$high_rent_value==0,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[3]],temp2)
iv_reg[[3]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.3 & temp$high_rent_value==1,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[4]],temp2)
iv_reg[[4]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.5 & temp$high_rent_value==0,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[5]],temp2)
iv_reg[[5]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.5 & temp$high_rent_value==1,]
  temp2['mpay_value'] <- predict(iv_reg_1st[[6]],temp2)
iv_reg[[6]] <- lm(iv_formula_2, data=temp2)

stargazer(iv_reg,type = output.type,no.space = TRUE,omit =omit,omit.labels = omit.labels,column.labels   = c("10% NE (rv<med)","10% NE (rv>med)","30% NE (rv<med)","30% NE (rv>med)", "50% NE (rv<med)", "50% NE (rv>med)"), omit.stat = omit.stat)

```

<a href="#top">Back to top</a>

## Negative Equity: IV Regressions (Split by mortgage-to-rent)
* Less than median: Loans from zip codes where median mortgage-to-rent is less than the median(mortgage-to-rent) in the same quarter
* Greater than median: Loans from zip codes where median mortgage-to-rent is greater than the median(mortgage-to-rent) in the same quarter
```{r tslsrent_ne_2nd_3}
temp <- loan_data[loan_data$ne_cutoff %in% c(-0.1,-0.3,-0.5),]

temp <- completeFun(temp,c("def","mpay_value","fico","ltv","dti","upb","current_housevalue","interest_diff","zip_year","zip_orgyear","mtg30us","age_t","high_mpay_rent"))

iv_reg_1st <- list()
iv_reg <- list()


  temp2 <- temp[temp$ne_cutoff==-0.1 & temp$high_mpay_rent==0,]
  iv_reg_1st[[1]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[1]],temp2)
iv_reg[[1]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.1 & temp$high_mpay_rent==1,]
  iv_reg_1st[[2]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[2]],temp2)
iv_reg[[2]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.3 & temp$high_mpay_rent==0,]
  iv_reg_1st[[3]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[3]],temp2)
iv_reg[[3]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.3 & temp$high_mpay_rent==1,]
  iv_reg_1st[[4]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[4]],temp2)
iv_reg[[4]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.5 & temp$high_mpay_rent==0,]
  iv_reg_1st[[5]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[5]],temp2)
iv_reg[[5]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==-0.5 & temp$high_mpay_rent==1,]
  iv_reg_1st[[6]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[6]],temp2)
iv_reg[[6]] <- lm(iv_formula_2, data=temp2)

stargazer(iv_reg,type = output.type,no.space = TRUE,omit =omit,omit.labels = omit.labels,column.labels   = c("10% NE (mr<med)","10% NE (mr>med)","30% NE (mr<med)","30% NE (mr>med)", "50% NE (mr<med)", "50% NE (mr>med)"), omit.stat = omit.stat)

```

<a href="#top">Back to top</a>

# Positive Equity: Default ~ Mortgage-to-value + rent-to-value

* Sample: Loan enters in to the sample after borrower reaching the particular positive equity threshold and the default indicates whether borrower defaulted in the following 12 months after reaching the negative equity threshold.
* mtg30us: 30-Year Fixed Rate Mortgage Average in the United States in the loan origination month
* interest_diff: interest rate - mtg30us
* rent-to-value : median rent/median value for the zip code in the negative equity quarter

## Positive Equity: First Stage
```{r tslsrent_pe_1st}
temp <- loan_data[loan_data$ne_cutoff %in% c(0.1,0.3,0.5,0.7,0.9),]
# temp <- temp[temp$org_year<=2008 & temp$org_year>=2004 & temp$ne_year<=2012,]
# temp <- temp[sample(nrow(temp),100000),]

temp <- completeFun(temp,c("def","mpay_value","fico","ltv","dti","upb","current_housevalue","interest_diff","zip_year","zip_orgyear","mtg30us","age_t"))


iv_reg_1st <- list()
iv_reg_1st[[1]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==0.1,])
iv_reg_1st[[2]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==0.3,])
iv_reg_1st[[3]] <- lm(as.formula(first_stage_formula), data=temp[temp$ne_cutoff==0.5,])


stargazer(iv_reg_1st,type = output.type,no.space = TRUE,omit =omit,omit.labels = omit.labels,column.labels   = c("10% PE","30% PE","50% PE"),omit.stat = omit.stat)
```

<a href="#top">Back to top</a>

## Positive Equity: IV Regressions

```{r tslsrent_pe_2nd}
iv_reg <- list()
iv_reg_1st <- list()
iv_reg[[1]] <- lm(iv_formula, data=temp[temp$ne_cutoff==0.1,])

  temp2 <- temp[temp$ne_cutoff==0.1,]
   iv_reg_1st[[1]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[1]])
iv_reg[[2]] <- lm(iv_formula, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.3,]
  iv_reg_1st[[2]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[2]])
iv_reg[[3]] <- lm(iv_formula, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.5,]
  iv_reg_1st[[3]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[3]])
iv_reg[[4]] <- lm(iv_formula, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.7,]
  iv_reg_1st[[4]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[4]])
iv_reg[[5]] <- lm(iv_formula, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.9,]
  iv_reg_1st[[5]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[5]])
iv_reg[[6]] <- lm(iv_formula, data=temp2)

stargazer(iv_reg,type = output.type,no.space = TRUE,omit =omit,omit.labels = omit.labels,column.labels   = c("OLS 10% PE","IV 10% PE","IV 30% PE","IV 50% PE","IV 70% PE","IV 90% PE"), omit.stat = omit.stat)
```

<a href="#top">Back to top</a>

## Positive Equity: IV Regressions (Split by rent-to-value)

* Less than median: Loans from zip codes where median rent-to-value is less than the median(rent-to-value) in the same quarter
* Greater than median: Loans from zip codes where median rent-to-value is greater than the median(rent-to-value) in the same quarter

```{r tslsrent_pe_2nd_2}

iv_reg_1st <- list()

iv_reg <- list()
  temp2 <- temp[temp$ne_cutoff==0.1 & temp$high_rent_value==0,]
  iv_reg_1st[[1]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[1]])
iv_reg[[1]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.1 & temp$high_rent_value==1,]
  iv_reg_1st[[2]] <- lm(as.formula(first_stage_formula),data = temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[2]])
iv_reg[[2]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.3 & temp$high_rent_value==0,]
  iv_reg_1st[[3]] <- lm(as.formula(first_stage_formula),data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[3]])
iv_reg[[3]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.3 & temp$high_rent_value==1,]
  iv_reg_1st[[4]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[4]])
iv_reg[[4]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.5 & temp$high_rent_value==0,]
  iv_reg_1st[[5]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[5]])
iv_reg[[5]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.5 & temp$high_rent_value==1,]
  iv_reg_1st[[6]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[6]])
iv_reg[[6]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.7 & temp$high_rent_value==0,]
  iv_reg_1st[[7]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[7]])
iv_reg[[7]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.7 & temp$high_rent_value==1,]
  iv_reg_1st[[8]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[8]])
iv_reg[[8]] <- lm(iv_formula_2, data=temp2)

stargazer(iv_reg,type = output.type,no.space = TRUE,omit =omit,omit.labels = omit.labels,column.labels   = c("10% PE (rv<med)","10% PE (rv>med)","30% PE (rv<med)","30% PE (rv>med)", "50% PE (rv<med)", "50% PE (rv>med)", "70% PE (rv<med)", "70% PE (rv>med)"), omit.stat = omit.stat)

```

<a href="#top">Back to top</a>

## Positive Equity: IV Regressions (Split by mortgage-to-rent)

* Less than median: Loans from zip codes where median mortgage-to-rent is less than the median(mortgage-to-rent) in the same quarter
* Greater than median: Loans from zip codes where median mortgage-to-rent is greater than the median(mortgage-to-rent) in the same quarter

```{r tslsrent_pe_2nd_3}

iv_reg_1st <- list()

iv_reg <- list()
  temp2 <- temp[temp$ne_cutoff==0.1 & temp$high_mpay_rent==0,]
  iv_reg_1st[[1]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[1]])
iv_reg[[1]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.1 & temp$high_mpay_rent==1,]
  iv_reg_1st[[2]] <- lm(as.formula(first_stage_formula),data = temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[2]])
iv_reg[[2]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.3 & temp$high_mpay_rent==0,]
  iv_reg_1st[[3]] <- lm(as.formula(first_stage_formula),data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[3]])
iv_reg[[3]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.3 & temp$high_mpay_rent==1,]
  iv_reg_1st[[4]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[4]])
iv_reg[[4]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.5 & temp$high_mpay_rent==0,]
  iv_reg_1st[[5]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[5]])
iv_reg[[5]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.5 & temp$high_mpay_rent==1,]
  iv_reg_1st[[6]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[6]])
iv_reg[[6]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.7 & temp$high_mpay_rent==0,]
  iv_reg_1st[[7]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[7]])
iv_reg[[7]] <- lm(iv_formula_2, data=temp2)

  temp2 <- temp[temp$ne_cutoff==0.7 & temp$high_mpay_rent==1,]
  iv_reg_1st[[8]] <- lm(as.formula(first_stage_formula), data=temp2)
  temp2['mpay_value'] <- predict(iv_reg_1st[[8]])
iv_reg[[8]] <- lm(iv_formula_2, data=temp2)

stargazer(iv_reg,type = output.type,no.space = TRUE,omit =omit,omit.labels = omit.labels,column.labels   = c("10% PE (mr<med)","10% PE (mr>med)","30% PE (mr<med)","30% PE (mr>med)", "50% PE (mr<med)", "50% PE (mr>med)", "70% PE (mr<med)", "70% PE (mr>med)"), omit.stat = omit.stat)

```

```{r stataconvert,include=FALSE}
# library(foreign)
# DF <- loan_data
# DF[sapply(DF, is.character)] <- lapply(DF[sapply(DF,is.character)],as.factor)
# write.dta(DF,file="loan_data.dta")
```



