---
title: "Strategic Default"
date: "October 2, 2016"
output: 
  html_document: 
    css: bodycss.css
    fig_width: 12
    fig_height: 6
---

Using zillow rent.

```{r warning=FALSE,message=FALSE}
rm(list=ls())
library(data.table)
library(zoo)
library(sqldf)
library(stargazer)
library(survival)
library(reshape2)
library(zipcode)
library(ggplot2)
library(sandwich)
file_path = "E:/strategic_default/"
setwd(file_path)
source("functions.R")
output.type = "html"

zillow_cox <- readRDS("cox_ne_0_Oct142016.rds")
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.1_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.2_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.3_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.4_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.5_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.6_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.7_Oct142016.rds"))

zillow_cox['rec_state']<-ifelse(zillow_cox$recourse==1,"rec","non-rec")

zip_forc_time <- readRDS("zip_forc_time.rds")
zillow_cox <- merge(zillow_cox,zip_forc_time,by="zip",all.x = TRUE)
mpay_table <- readRDS("mpay_table.rds")

```

### Descriptive Statistics
```{r warning=FALSE,results='asis'}
zillow_cox$year <- as.integer(zillow_cox$year)
stargazer(
  zillow_cox[zillow_cox$ne_cutoff==0, c("upb","fico","ltv","interestrate","adjusted_mpay_rent","mpay","year","median_forc_time")], type = output.type, 
  summary.stat = c("mean", "sd","min", "p25", "median", "p75", "max","n")
)

```


### Most borrowers do not default when faced with negative equity
```{r results='asis'}
default_pct <- ddply(zillow_cox,.(ne_cutoff),summarise,default_rate = mean(def,na.rm = TRUE))
default_pct <- default_pct[order(-default_pct$ne_cutoff),]
ggplot(default_pct, aes(x = factor(ne_cutoff), y = default_rate)) + geom_bar(stat = "identity")+ylab("Default Rate")+xlab("Negative Equity")
```

### Mortgage payment vs Rent comparison
Compares the mortgage payment of a typical loan originated in 2005 in each zipcode to the zillow single family rent for each zip code from 2010 to 2014
```{r}
# allloans <- readRDS("CombinedFreddieFannieAct_Nov18.rds")
# allloans['org_year']<-format(allloans$orgdate,"%Y")
# allloans <- allloans[allloans$org_year==2005,]
# gc(reset = TRUE)
# mpay_table <- ddply(allloans,.(zip),summarise,upb=median(upb,na.rm = TRUE),ltv=median(ltv,na.rm = TRUE),interestrate=median(interestrate,na.rm = TRUE))
# saveRDS(mpay_table,file="mpay_table.rds")
mpay_table['mpay']<-apply(mpay_table[,c('upb','interestrate')],1,function(x) mortgage(x['upb'],x['interestrate'],360))
zillow_zip <- ddply(zillow_cox,.(zip),summarise,rent = max(rent,na.rm = TRUE))
zillow_zip <- merge(mpay_table,zillow_zip,by=c("zip"))
zillow_zip['mpay_minus_rent'] <- zillow_zip$mpay - zillow_zip$rent
zillow_zip <- zillow_zip[zillow_zip$mpay_minus_rent<quantile(zillow_zip$mpay_minus_rent,0.99) & zillow_zip$mpay_minus_rent>quantile(zillow_zip$mpay_minus_rent,0.01),]
summary(zillow_zip$mpay_minus_rent)
qplot(zillow_zip$mpay_minus_rent, geom="histogram", main = "Histogram for Mortgage Payment - Rent", 
      xlab = "Mortgage Payment - Rent",  
      alpha=I(1),
      xlim=c(-1500,200))
```


### Main Cox Model: All variables
```{r results='asis'}

cox_formula_1 <- as.formula("Surv(duration,def)~adjusted_mpay_rent+median_forc_time+age_t+factor(ne_year)+factor(recourse)+factor(ne_cutoff)")

cox_formula_2 <- as.formula("Surv(duration,def)~adjusted_mpay_rent+median_forc_time+age_t+age_t*age_t+del_unemp+factor(ne_year)+fico+ltv+dti+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(occupancy)+factor(loanpurpose)+factor(recourse)+factor(ne_cutoff)")

cox <-list()
cox[[1]] <- coxph(cox_formula_1,data = zillow_cox)
cox[[2]] <- coxph(cox_formula_2,data = zillow_cox)

stargazer(cox, omit = c("ne_year","loanpurpose","occupancy","ne_cutoff"),omit.labels = c("Year","","",""),type = output.type,no.space=TRUE)
```


### Hazard Contribution
```{r}

# adjusted_mpay_rent polynomial
cox_formula <- as.formula("Surv(duration,def)~adjusted_mpay_rent+I(adjusted_mpay_rent^2)+I(adjusted_mpay_rent^3)+I(adjusted_mpay_rent^4)+I(adjusted_mpay_rent^5)+median_forc_time+age_t+I(median_forc_time^2)+I(median_forc_time^3)+I(median_forc_time^4)+I(median_forc_time^5)+age_t*age_t+del_unemp+factor(ne_year)+fico+ltv+dti+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+log(upb)+factor(occupancy)+factor(loanpurpose)+factor(recourse)+factor(ne_cutoff)")

cox <- coxph(cox_formula,data = zillow_cox)

x = (40:160)/100
y = (summary(cox)$coefficients)[1,1] * x +(summary(cox)$coefficients)[2,1] * x^2 +(summary(cox)$coefficients)[3,1] * x^3 +(summary(cox)$coefficients)[4,1] * x^4 +(summary(cox)$coefficients)[5,1] * x^5
plot_data <- as.data.frame(cbind(x,y))

ggplot(plot_data,aes(x=x,y=y))+geom_point()+
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+labs(x = "adjusted_mpay_rent",y="proportional hazard contribution")


# median_forc_time polynomial
cox_formula <- as.formula("Surv(duration,def)~median_forc_time+I(median_forc_time^2)+I(median_forc_time^3)+I(median_forc_time^4)+I(median_forc_time^5)+adjusted_mpay_rent+I(adjusted_mpay_rent^2)+I(adjusted_mpay_rent^3)+I(adjusted_mpay_rent^4)+I(adjusted_mpay_rent^5)+age_t+age_t*age_t+del_unemp+factor(state)+factor(ne_year)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(occupancy)+factor(loanpurpose)+factor(recourse)+factor(ne_cutoff)")

cox <- coxph(cox_formula,data = zillow_cox)

x = seq(from=11, to=16,by=0.1)
y = (summary(cox)$coefficients)[1,1] * x +(summary(cox)$coefficients)[2,1] * x^2 +(summary(cox)$coefficients)[3,1] * x^3 +(summary(cox)$coefficients)[4,1] * x^4 +(summary(cox)$coefficients)[5,1] * x^5
plot_data <- as.data.frame(cbind(x,y))

ggplot(plot_data,aes(x=x,y=y))+geom_point()+
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+labs(x = "median_forc_time",y="proportional hazard contribution")


```

### Cox regression (default,duration)~adjusted_mpay_rent (2)
```{r results='asis',warning=FALSE}
ne_cuts <- c(0,-0.1,-0.2,-0.3,-0.4,-0.5,-0.6,-0.7)
i=1

cox_formula <- as.formula("Surv(duration,def)~adjusted_mpay_rent+poly(mean_forc_time,5)+age_t+age_t*age_t+del_unemp+factor(ne_year)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(occupancy)+factor(loanpurpose)+factor(rec_state)")

cox<-list()
cox_coefficients <- NULL
for(cut in ne_cuts) {
  cat(i,cut,"\n")
  cox[[i]] <- coxph(cox_formula,data = zillow_cox[zillow_cox$ne_cutoff==cut & !is.na(zillow_cox$median_forc_time),])
  
  cox_coefficients <- rbind(cox_coefficients, c(cut,summary(cox[[i]])$coefficients[1,1],summary(cox[[i]])$coefficients[1,3],(exp(as.numeric(cox[[i]]$coefficients[1]))-1)*sd(zillow_cox[zillow_cox$ne_cutoff==cut,]$adjusted_mpay_rent)))

  
  i=i+1
}

#stargazer(cox, omit = c("ne_year","loanpurpose","occupancy"),omit.labels = c("Year","",""),type = output.type,no.space=TRUE)
cox_coefficients <- as.data.frame(cox_coefficients)
names(cox_coefficients) <- c("NE Cutoff","Coefficient","Std. Error","Marginal Effect")
stargazer(cox_coefficients,summary = FALSE,type = output.type)

```


### Zipcode level regression: default % ~ median mpay_rent

```{r warning=FALSE}
zillow_cox['ne_qt']<- as.yearqtr(zillow_cox$reportingperiod_t)
median_mpay_zillow <- ddply(zillow_cox,.(ne_qt,ne_cutoff,recourse,state,zip),summarise,
                     def=mean(def,na.rm = TRUE),
                     adjusted_mpay_rent = mean(adjusted_mpay_rent,na.rm = TRUE),
                     fico = mean(fico,na.rm = TRUE),
                     ltv = mean(ltv,na.rm = TRUE),
                     dti = mean(dti,na.rm = TRUE),
                     ne_year = mean(ne_year,na.rm = TRUE),
                     upb = mean(upb,na.rm = TRUE),
                     age_t = mean(age_t, na.rm = TRUE),
                     medhhincome = mean(medhhincome,na.rm = TRUE),
                     medhouseage = mean(medhouseage,na.rm = TRUE),
                     vacantpct = mean(vacantpct,na.rm = TRUE),
                     median_forc_time = mean(median_forc_time,na.rm = TRUE),
                     bachelorpct = mean(bachelorpct,na.rm = TRUE),
                     withsalarypct = mean(withsalarypct,na.rm = TRUE),
                     obs = length(loanid))

median_mpay_zillow2 <- ddply(zillow_cox,.(ne_cutoff,recourse,state),summarise,
                     def=mean(def,na.rm = TRUE),
                     adjusted_mpay_rent = mean(adjusted_mpay_rent,na.rm = TRUE),
                     fico = mean(fico,na.rm = TRUE),
                     ltv = mean(ltv,na.rm = TRUE),
                     dti = mean(dti,na.rm = TRUE),
                     ne_year = mean(ne_year,na.rm = TRUE),
                     upb = mean(upb,na.rm = TRUE),
                     age_t = mean(age_t, na.rm = TRUE),
                     medhhincome = mean(medhhincome,na.rm = TRUE),
                     medhouseage = mean(medhouseage,na.rm = TRUE),
                     vacantpct = mean(vacantpct,na.rm = TRUE),
                     median_forc_time = mean(median_forc_time,na.rm = TRUE),
                     bachelorpct = mean(bachelorpct,na.rm = TRUE),
                     withsalarypct = mean(withsalarypct,na.rm = TRUE),
                     obs = length(loanid))

summary(median_mpay_zillow2$def)
summary(median_mpay_zillow2$adjusted_mpay_rent)

```

```{r  results='asis',warning=FALSE}
ols <- list()
se <- list()
marginal_effects <- list()

ols_formula <- as.formula("def~adjusted_mpay_rent+median_forc_time+factor(recourse)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(state)+factor(ne_qt)+age_t")

#------------------
ols[[1]]<-lm(ols_formula,data=median_mpay_zillow[median_mpay_zillow$ne_cutoff==0,])
se[[1]] <- sqrt(diag(vcovHAC(ols[[1]])))
marginal_effects[[1]] <- summary(ols[[1]])$coefficients[2]*sd(median_mpay_zillow[median_mpay_zillow$ne_cutoff==0,]$adjusted_mpay_rent)
#------------------
ols[[2]]<-lm(ols_formula,data=median_mpay_zillow[median_mpay_zillow$ne_cutoff==-0.1,])
se[[2]] <- sqrt(diag(vcovHAC(ols[[2]])))
marginal_effects[[2]] <- summary(ols[[2]])$coefficients[2]*sd(median_mpay_zillow[median_mpay_zillow$ne_cutoff==-0.1,]$adjusted_mpay_rent)
#------------------
ols[[3]]<-lm(ols_formula,data=median_mpay_zillow[median_mpay_zillow$ne_cutoff==-0.2,])

se[[3]] <- sqrt(diag(vcovHAC(ols[[3]])))
marginal_effects[[3]] <- summary(ols[[3]])$coefficients[2]*sd(median_mpay_zillow[median_mpay_zillow$ne_cutoff==-0.2,]$adjusted_mpay_rent)
#------------------
ols[[4]]<-lm(def~adjusted_mpay_rent+median_forc_time+factor(recourse)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(state)+factor(ne_qt)+age_t+factor(ne_cutoff),data=median_mpay_zillow)
#se[[4]] <- sqrt(diag(vcovHAC(ols[[4]])))
marginal_effects[[4]] <- summary(ols[[4]])$coefficients[2]*sd(median_mpay_zillow$adjusted_mpay_rent)
#------------------
stargazer(ols, omit = c("state","ne_qt","ne_cutoff"), column.labels=c('NE 0','NE -10%','NE -20%','All NE'), omit.labels = c("State","Time","Negative Equity"), type=output.type,no.space=TRUE,se=se)

marginal_effects
```

## Delete Later
### Loan-month default
```{r results='asis'}
# ne_cuts <- c(0,-0.1,-0.2,-0.3,-0.4,-0.5,-0.6,-0.7)
# i=1
# logits <- list()
# 
# 
# for(cut in ne_cuts) {
#   gc()
#   fn_loanmonth <- paste(file_path,"loanmonth_ne_",cut,"_Oct142016.rds",sep="")
#   loanmonth <- readRDS(fn_loanmonth)
#   gc()
#   logits[[i]] <- glm(factor(delstatus_t)~adjusted_mpay_rent+interestrate+fico+ltv+dti+
#                        poly(age_since_ne,5)+factor(state)+factor(ne_year),
#                      family='binomial',data=loanmonth,maxit=100)
#   print(summary(logits[[i]])$coefficients[2])
#   i=i+1
#   gc()
# }

```


### Univariate Evidence Default ~ mpay_rent
```{r warning=FALSE,results='asis'}

# quantiles <- ddply(zillow_cox,.(ne_cutoff,recourse),summarise,
#                    p20 = quantile(adjusted_mpay_rent,0.2,na.rm = TRUE),
#                    p40 = quantile(adjusted_mpay_rent,0.4,na.rm = TRUE),
#                    p60 = quantile(adjusted_mpay_rent,0.6,na.rm = TRUE),
#                    p80 = quantile(adjusted_mpay_rent,0.8,na.rm = TRUE),
#                    p100 = quantile(adjusted_mpay_rent,1,na.rm = TRUE))
# 
# zillow_cox <- merge(zillow_cox,quantiles,by=c("ne_cutoff","recourse"))
# 
# zillow_cox['adj_mpay_rent_q']<-ifelse(zillow_cox$adjusted_mpay_rent<= zillow_cox$p20,1,
#                                    ifelse(zillow_cox$adjusted_mpay_rent<= zillow_cox$p40,2,
#                                       ifelse(zillow_cox$adjusted_mpay_rent<= zillow_cox$p60,3,
#                                          ifelse(zillow_cox$adjusted_mpay_rent<= zillow_cox$p80,4,5))))
# 
# mpay_defpct <- ddply(zillow_cox, .(ne_cutoff,adj_mpay_rent_q),summarise,default=mean(def))
# mpay_defpct<-melt(mpay_defpct,id=c( "ne_cutoff","adj_mpay_rent_q"), measured=c("default"))
# mpay_defpct<-dcast(mpay_defpct, adj_mpay_rent_q~ne_cutoff)
# 
# stargazer(mpay_defpct, type = output.type,summary = FALSE)
# 
# 
# cox_univariate <- coxph(Surv(duration,def)~poly(adjusted_mpay_rent,5),data = zillow_cox[zillow_cox$recourse==0 & zillow_cox$ne_cutoff==-0.2,])

```

### Cox regression (default,duration)~adjusted_mpay_rent
```{r results='asis',warning=FALSE}
# ne_cuts <- c(0,-0.1,-0.2,-0.3,-0.4,-0.5,-0.6,-0.7)
# i=1
# 
# cox_formula <- as.formula("Surv(duration,def)~adjusted_mpay_rent+age_t+age_t*age_t+del_unemp+factor(state)+factor(ne_year)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(occupancy)+factor(loanpurpose)")
# 
# cox<-list()
# cox_coefficients <- NULL
# for(cut in ne_cuts) {
#   cox[[i]] <- coxph(cox_formula,data = zillow_cox[zillow_cox$recourse==0 & zillow_cox$ne_cutoff==cut,])
#   cox_coefficients <- rbind(cox_coefficients, c(cut,0,summary(cox[[i]])$coefficients[1,1],summary(cox[[i]])$coefficients[1,3],(exp(as.numeric(cox[[i]]$coefficients[1]))-1)*sd(zillow_cox[zillow_cox$recourse==0 & zillow_cox$ne_cutoff==cut,]$adjusted_mpay_rent)))
#   i=i+1
#   cox[[i]] <- coxph(cox_formula,data = zillow_cox[zillow_cox$recourse==1 & zillow_cox$ne_cutoff==cut,])
#   cox_coefficients <- rbind(cox_coefficients, c(cut,1,summary(cox[[i]])$coefficients[1,1],summary(cox[[i]])$coefficients[1,3],(exp(as.numeric(cox[[i]]$coefficients[1]))-1)*sd(zillow_cox[zillow_cox$recourse==1 & zillow_cox$ne_cutoff==cut,]$adjusted_mpay_rent)))
#   i=i+1
# }
# 
# cox_coefficients <- as.data.frame(cox_coefficients)
# names(cox_coefficients) <- c("NE Cutoff","Recourse","Coefficient","Std. Error","Marginal Effect")
# stargazer(cox[[5]],cox[[6]],cox[[9]],cox[[10]],column.labels=c('Non-Recourse-20','Recourse-20','Non-Recourse-40','Recourse-40'), omit = c("state","ne_year","loanpurpose","occupancy"),omit.labels = c("State","Year","",""),type = output.type,no.space=TRUE)
# 
# cox_coefficients <- cox_coefficients[order(cox_coefficients$Recourse),]
# stargazer(cox_coefficients,summary = FALSE,type = output.type)
```

