---
title: "Strategic Default"
date: "October 2, 2016"
output: 
  html_document: 
    css: bodycss.css
    fig_width: 12
    fig_height: 6
---

Using zillow rent.

```{r warning=FALSE,message=FALSE}
rm(list=ls())
library(data.table)
library(zoo)
library(sqldf)
library(stargazer)
library(survival)
library(reshape2)
library(zipcode)
library(ggplot2)
library(sandwich)
file_path = "E:/strategic_default/"
setwd(file_path)
source("functions.R")
output.type = "text"

zillow_cox <- readRDS("cox_ne_0_Oct142016.rds")
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.1_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.2_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.3_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.4_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.5_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.6_Oct142016.rds"))
zillow_cox <- rbind(zillow_cox,readRDS("cox_ne_-0.7_Oct142016.rds"))

zillow_cox['rec_state']<-ifelse(zillow_cox$recourse==1,"rec","non-rec")

zip_forc_time <- readRDS("zip_forc_time.rds")
zillow_cox <- merge(zillow_cox,zip_forc_time,by="zip",all.x = TRUE)
mpay_table <- readRDS("mpay_table.rds")

```

### Descriptive Statistics
```{r warning=FALSE,results='asis'}
zillow_cox$year <- as.integer(zillow_cox$year)
stargazer(
  zillow_cox[zillow_cox$ne_cutoff==0, c("upb","fico","ltv","interestrate","adjusted_mpay_rent","mpay","year","median_forc_time")], type = output.type, 
  summary.stat = c("mean", "sd","min", "p25", "median", "p75", "max","n")
)

```


### mpay_rent map
```{r}
require(RCurl)
require(xlsx)
require(zipcode)
require(ggmap)
require("ggplot2")

zip_mpay_rent <- ddply(zillow_cox,.(zip),summarise,
                     adjusted_mpay_rent=mean(adjusted_mpay_rent,na.rm = TRUE))
head(zip_mpay_rent)

data(zipcode)
names(zipcode)[1] = "zip5"
zipcode$zip5 <- as.numeric(zipcode$zip5)
zipcode['zip']<-floor(zipcode$zip5/100)*100
zip_mpay_rent <- merge(zipcode, zip_mpay_rent, by='zip')

map<-get_map(location='united states', zoom=4, maptype = "terrain",
             source='google',color='color')

ggmap(map) + geom_point(
        aes(x=longitude, y=latitude, show_guide = TRUE, colour=adjusted_mpay_rent), 
        data=zip_mpay_rent, alpha=.8, na.rm = T)  + 
        scale_color_gradient(low="yellow", high="blue")

```


### Most borrowers do not default when faced with negative equity
```{r results='asis'}
default_pct <- ddply(zillow_cox,.(ne_cutoff),summarise,default_rate = mean(def,na.rm = TRUE))
default_pct <- default_pct[order(-default_pct$ne_cutoff),]
ggplot(default_pct, aes(x = factor(ne_cutoff), y = default_rate)) + geom_bar(stat = "identity")+ylab("Default Rate")+xlab("Negative Equity")
```

### Mortgage payment vs Rent comparison
Compares the mortgage payment of a typical loan originated in 2005 in each zipcode to the zillow single family rent for each zip code from 2010 to 2014
```{r}
# allloans <- readRDS("CombinedFreddieFannieAct_Nov18.rds")
# allloans['org_year']<-format(allloans$orgdate,"%Y")
# allloans <- allloans[allloans$org_year==2005,]
# gc(reset = TRUE)
# mpay_table <- ddply(allloans,.(zip),summarise,upb=median(upb,na.rm = TRUE),ltv=median(ltv,na.rm = TRUE),interestrate=median(interestrate,na.rm = TRUE))
# saveRDS(mpay_table,file="mpay_table.rds")
mpay_table['mpay']<-apply(mpay_table[,c('upb','interestrate')],1,function(x) mortgage(x['upb'],x['interestrate'],360))
zillow_zip <- ddply(zillow_cox,.(zip),summarise,rent = max(rent,na.rm = TRUE))
zillow_zip <- merge(mpay_table,zillow_zip,by=c("zip"))
zillow_zip['mpay_minus_rent'] <- zillow_zip$mpay - zillow_zip$rent
zillow_zip <- zillow_zip[zillow_zip$mpay_minus_rent<quantile(zillow_zip$mpay_minus_rent,0.99) & zillow_zip$mpay_minus_rent>quantile(zillow_zip$mpay_minus_rent,0.01),]
summary(zillow_zip$mpay_minus_rent)
qplot(zillow_zip$mpay_minus_rent, geom="histogram", main = "Histogram for Mortgage Payment - Rent", 
      xlab = "Mortgage Payment - Rent",  
      alpha=I(1),
      xlim=c(-1500,200))
```


### Main Cox Model: All variables
```{r results='asis'}

cox_formula_1 <- as.formula("Surv(duration,def)~adjusted_mpay_rent+median_forc_time+age_t+factor(ne_year)+factor(recourse)+factor(ne_cutoff)+log(upb)/ltv")

cox_formula_2 <- as.formula("Surv(duration,def)~adjusted_mpay_rent+median_forc_time+age_t+age_t*age_t+del_unemp+factor(ne_year)+fico+ltv+dti+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(occupancy)+factor(loanpurpose)+factor(recourse)+factor(ne_cutoff)+log(upb)/ltv")

cox <-list()
cox[[1]] <- coxph(cox_formula_1,data = zillow_cox)
cox[[2]] <- coxph(cox_formula_2,data = zillow_cox)

stargazer(cox, omit = c("ne_year","loanpurpose","occupancy","ne_cutoff"),omit.labels = c("Year","","",""),type = output.type,no.space=TRUE)
```


### Hazard Contribution
```{r}

# adjusted_mpay_rent polynomial
cox_formula <- as.formula("Surv(duration,def)~adjusted_mpay_rent+I(adjusted_mpay_rent^2)+I(adjusted_mpay_rent^3)+I(adjusted_mpay_rent^4)+I(adjusted_mpay_rent^5)+median_forc_time+age_t+I(median_forc_time^2)+I(median_forc_time^3)+I(median_forc_time^4)+I(median_forc_time^5)+age_t*age_t+del_unemp+factor(ne_year)+fico+ltv+dti+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+log(upb)+factor(occupancy)+factor(loanpurpose)+factor(recourse)+factor(ne_cutoff)+log(upb)/ltv")

cox <- coxph(cox_formula,data = zillow_cox)

x = (40:160)/100
y = (summary(cox)$coefficients)[1,1] * x +(summary(cox)$coefficients)[2,1] * x^2 +(summary(cox)$coefficients)[3,1] * x^3 +(summary(cox)$coefficients)[4,1] * x^4 +(summary(cox)$coefficients)[5,1] * x^5
plot_data <- as.data.frame(cbind(x,y))

ggplot(plot_data,aes(x=x,y=y))+geom_point()+
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+labs(x = "adjusted_mpay_rent",y="proportional hazard contribution")


# median_forc_time polynomial
cox_formula <- as.formula("Surv(duration,def)~median_forc_time+I(median_forc_time^2)+I(median_forc_time^3)+I(median_forc_time^4)+I(median_forc_time^5)+adjusted_mpay_rent+I(adjusted_mpay_rent^2)+I(adjusted_mpay_rent^3)+I(adjusted_mpay_rent^4)+I(adjusted_mpay_rent^5)+age_t+age_t*age_t+del_unemp+factor(state)+factor(ne_year)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(occupancy)+factor(loanpurpose)+factor(recourse)+factor(ne_cutoff)+log(upb)/ltv")

cox <- coxph(cox_formula,data = zillow_cox)

x = seq(from=11, to=16,by=0.1)
y = (summary(cox)$coefficients)[1,1] * x +(summary(cox)$coefficients)[2,1] * x^2 +(summary(cox)$coefficients)[3,1] * x^3 +(summary(cox)$coefficients)[4,1] * x^4 +(summary(cox)$coefficients)[5,1] * x^5
plot_data <- as.data.frame(cbind(x,y))

ggplot(plot_data,aes(x=x,y=y))+geom_point()+
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+labs(x = "median_forc_time",y="proportional hazard contribution")


```

### Cox regression (default,duration)~adjusted_mpay_rent (2)
```{r results='asis',warning=FALSE}
ne_cuts <- c(0,-0.1,-0.2,-0.3,-0.4,-0.5,-0.6,-0.7)
i=1

cox_formula <- as.formula("Surv(duration,def)~adjusted_mpay_rent+poly(mean_forc_time,5)+age_t+age_t*age_t+del_unemp+factor(ne_year)+fico+ltv+dti+log(upb)+log(medhhincome)+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(occupancy)+factor(loanpurpose)+factor(rec_state)+log(upb)/ltv")

cox<-list()
cox_coefficients <- NULL
for(cut in ne_cuts) {
  cox[[i]] <- coxph(cox_formula,data = zillow_cox[zillow_cox$ne_cutoff==cut & !is.na(zillow_cox$median_forc_time),])
  
  cox_coefficients <- rbind(cox_coefficients, c(cut,summary(cox[[i]])$coefficients[1,1],summary(cox[[i]])$coefficients[1,3],(exp(as.numeric(cox[[i]]$coefficients[1]))-1)*sd(zillow_cox[zillow_cox$ne_cutoff==cut,]$adjusted_mpay_rent)))

  
  i=i+1
}

#stargazer(cox, omit = c("ne_year","loanpurpose","occupancy"),omit.labels = c("Year","",""),type = output.type,no.space=TRUE)
cox_coefficients <- as.data.frame(cox_coefficients)
names(cox_coefficients) <- c("NE Cutoff","Coefficient","Std. Error","Marginal Effect")
stargazer(cox_coefficients,summary = FALSE,type = output.type)

```


### Zipcode level regression: default % ~ median mpay_rent

```{r warning=FALSE}
zillow_cox['ne_qt']<- as.yearqtr(zillow_cox$reportingperiod_t)
median_mpay_zillow <- ddply(zillow_cox,.(ne_qt,ne_cutoff,recourse,state,zip),summarise,
                     def=mean(def,na.rm = TRUE),
                     adjusted_mpay_rent = mean(adjusted_mpay_rent,na.rm = TRUE),
                     fico = mean(fico,na.rm = TRUE),
                     ltv = mean(ltv,na.rm = TRUE),
                     dti = mean(dti,na.rm = TRUE),
                     ne_year = mean(ne_year,na.rm = TRUE),
                     upb = mean(upb,na.rm = TRUE),
                     age_t = mean(age_t, na.rm = TRUE),
                     medhhincome = mean(medhhincome,na.rm = TRUE),
                     medhouseage = mean(medhouseage,na.rm = TRUE),
                     vacantpct = mean(vacantpct,na.rm = TRUE),
                     median_forc_time = mean(median_forc_time,na.rm = TRUE),
                     bachelorpct = mean(bachelorpct,na.rm = TRUE),
                     withsalarypct = mean(withsalarypct,na.rm = TRUE),
                     obs = length(loanid))

median_mpay_zillow2 <- ddply(zillow_cox,.(ne_cutoff,recourse,state),summarise,
                     def=mean(def,na.rm = TRUE),
                     adjusted_mpay_rent = mean(adjusted_mpay_rent,na.rm = TRUE),
                     fico = mean(fico,na.rm = TRUE),
                     ltv = mean(ltv,na.rm = TRUE),
                     dti = mean(dti,na.rm = TRUE),
                     ne_year = mean(ne_year,na.rm = TRUE),
                     upb = mean(upb,na.rm = TRUE),
                     age_t = mean(age_t, na.rm = TRUE),
                     medhhincome = mean(medhhincome,na.rm = TRUE),
                     medhouseage = mean(medhouseage,na.rm = TRUE),
                     vacantpct = mean(vacantpct,na.rm = TRUE),
                     median_forc_time = mean(median_forc_time,na.rm = TRUE),
                     bachelorpct = mean(bachelorpct,na.rm = TRUE),
                     withsalarypct = mean(withsalarypct,na.rm = TRUE),
                     obs = length(loanid))

summary(median_mpay_zillow2$def)
summary(median_mpay_zillow2$adjusted_mpay_rent)

```

```{r  results='asis',warning=FALSE}
ols <- list()
se <- list()
marginal_effects <- list()

ols_formula <- as.formula("def~adjusted_mpay_rent+median_forc_time+factor(recourse)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(state)+factor(ne_qt)+age_t+log(upb)/ltv")

#------------------
ols[[1]]<-lm(ols_formula,data=median_mpay_zillow[median_mpay_zillow$ne_cutoff==0,])
se[[1]] <- sqrt(diag(vcovHAC(ols[[1]])))
marginal_effects[[1]] <- summary(ols[[1]])$coefficients[2]*sd(median_mpay_zillow[median_mpay_zillow$ne_cutoff==0,]$adjusted_mpay_rent)
#------------------
ols[[2]]<-lm(ols_formula,data=median_mpay_zillow[median_mpay_zillow$ne_cutoff==-0.1,])
se[[2]] <- sqrt(diag(vcovHAC(ols[[2]])))
marginal_effects[[2]] <- summary(ols[[2]])$coefficients[2]*sd(median_mpay_zillow[median_mpay_zillow$ne_cutoff==-0.1,]$adjusted_mpay_rent)
#------------------
ols[[3]]<-lm(ols_formula,data=median_mpay_zillow[median_mpay_zillow$ne_cutoff==-0.2,])

se[[3]] <- sqrt(diag(vcovHAC(ols[[3]])))
marginal_effects[[3]] <- summary(ols[[3]])$coefficients[2]*sd(median_mpay_zillow[median_mpay_zillow$ne_cutoff==-0.2,]$adjusted_mpay_rent)
#------------------
ols[[4]]<-lm(def~adjusted_mpay_rent+median_forc_time+factor(recourse)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+factor(state)+factor(ne_qt)+age_t+factor(ne_cutoff),data=median_mpay_zillow)
#se[[4]] <- sqrt(diag(vcovHAC(ols[[4]])))
marginal_effects[[4]] <- summary(ols[[4]])$coefficients[2]*sd(median_mpay_zillow$adjusted_mpay_rent)
#------------------
stargazer(ols, omit = c("state","ne_qt","ne_cutoff"), column.labels=c('NE 0','NE -10%','NE -20%','All NE'), omit.labels = c("State","Time","Negative Equity"), type=output.type,no.space=TRUE,se=se)

marginal_effects
```


### Prepayment Loan-month
```{r results='asis',warning=FALSE}
deldata_zillow <- readRDS(paste(file_path,"prepay_loanmonth_zillow.rds",sep=""))

gc(reset = TRUE)

logit <- list()
logit[[1]] <- glm(prepaid~mpay_zillow_rent+interestrate+current_equity+I(age_t)+I(age_t^2)+I(age_t^3)+I(age_t^4)+I(age_t^5),family='binomial',data=deldata_zillow,maxit=100)
logit[[2]] <- glm(prepaid~mpay_zillow_rent+interestrate+current_equity+I(age_t)+I(age_t^2)+I(age_t^3)+I(age_t^4)+I(age_t^5),family='binomial',data=deldata_zillow[deldata_zillow$current_equity>=0,],maxit=100)
logit[[3]] <- glm(prepaid~mpay_zillow_rent+interestrate+current_equity+I(age_t)+I(age_t^2)+I(age_t^3)+I(age_t^4)+I(age_t^5)+fico+ltv+dti+log(upb)+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct,family='binomial',data=deldata_zillow[deldata_zillow$current_equity>=0,],maxit=100)

stargazer(logit,type=output.type,no.space=TRUE)

rm(deldata_zillow)
gc()

```



### Zip code level sold pct and inventory number
Percentage Houses Sold in each 3 digit zip code and for sale inventory in each zip code
(Using the median age to calculate mpay)
```{r results='asis',warning=FALSE}

decile_data <- readRDS(paste(file_path,"decile_stats_merged.rds",sep=""))
decile_data['new_default_pct']<- decile_data$freddie_no_of_defaults*100/decile_data$freddie_no_of_loans
decile_data$ne_default_pct=decile_data$ne_default_pct*100

stargazer(
  decile_data[decile_data$age_decile==5, c("sold_pct","ne_default_pct","inventory_value","mpay_rent","mpay","rent")], type = output.type, 
  summary.stat = c("mean", "sd","min", "p25", "median", "p75", "max")
)

ols <- list()
ols[[1]] <- lm(sold_pct~mpay_rent+fico+ltv+dti+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+age+factor(state)+factor(current_year),data=decile_data[decile_data$age_decile==5,])
ols[[2]] <- lm(inventory_value~mpay_rent+fico+ltv+dti+medhhincome+medhouseage+vacantpct+bachelorpct+withsalarypct+age+factor(state)+factor(current_year),data=decile_data[decile_data$age_decile==5,])

stargazer(ols, omit = c("state","current_year"),omit.labels = c("State","Year"),type=output.type,no.space=TRUE)

```